Subject-level developmental relationships

### 2024-10-04 1_get_data.qmd
refactor

```{r}
library(here)
source(here("helper","common.R"))
source(here("helper","rt_helper.R"))
```

@@

```{r}
d_sub <- d_trial |>
  group_by(dataset_name, subject_id, administration_id, age) |>
  summarise(across(c(contains("_window"),"rt","log_rt"), ~mean(.x, na.rm=TRUE)))
```

Add CDIs.

>>>>

```{r}
library(here)
source(here("helper","common.R"))
source(here("helper","rt_helper.R"))

# min number of trials
ACC_MIN_TRIAL_CUTOFF <- 4
RT_MIN_TRIAL_CUTOFF <- 2
```

@@

```{r}
d_sub <- d_trial |>
  group_by(dataset_name, subject_id, administration_id, age) |>
  summarise(n_trials = sum(!is.na(long_window_accuracy)), 
            n_trials_rt = sum(!is.na(rt)),
            rt_var = sd(rt, na.rm=TRUE), 
            log_rt_var = sd(log_rt, na.rm=TRUE), 
            short_window_acc_var = sd(short_window_accuracy, na.rm=TRUE),
            long_window_acc_var = sd(long_window_accuracy, na.rm=TRUE),
            across(c(contains("_window"),"rt","log_rt"), ~mean(.x, na.rm=TRUE)))
```

```{r}
ggplot(d_sub, aes(x = n_trials, fill = n_trials < 4)) + 
  geom_histogram(binwidth = 1) + 
  facet_wrap(~dataset_name, scales = "free_y")
```
```{r}
ggplot(d_sub, aes(x = n_trials_rt, fill = n_trials_rt < 2)) + 
  geom_histogram(binwidth = 1) + 
  facet_wrap(~dataset_name, scales = "free_y")
```

We impose the arbitrary cutoff that we need more than 4 accuracies and more than 2 RTs to analyze a participant.

```{r}
d_sub <- d_sub |>
  mutate(across(contains("log_rt"), ~ifelse(n_trials_rt < RT_MIN_TRIAL_CUTOFF, NA, .x)),
         across(contains("window"), ~ifelse(n_trials < ACC_MIN_TRIAL_CUTOFF, NA, .x)))

# Add CDI

### 2024-10-04 2_developmental_analysis.qmd
up to date for call
refactor

@@

```{r}
ggplot(d_sub_agevar, aes(x = age, y = short_window_accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

@@

kable(acc_mods_sub_lmer_summary, digits = 4)
```

@@

# Variability

```{r}
d_var <- d_trial |>
  group_by(dataset_name, administration_id, age) |>
  summarise (`SD(log RT)` = sd(log_rt, na.rm=TRUE), 
             `SD(accuracy)` = sd(long_window_elogit, na.rm=TRUE)) |>
  pivot_longer(contains("sd"), 
               values_to = "sd", names_to = "measure")

v <- ggplot(d_var, aes(x = age, y = sd)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_var |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name),
              aes(group = dataset_name, col = dataset_name), geom="line",
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  scale_color_discrete(guide = FALSE) + 
  scale_x_log10() +
  facet_wrap(~measure, scales = "free_y")

v 
ggsave(v, filename = "figures/var.png", width = 8, height = 3.5)

```
# Pub stuff

stats

```{r}
d_sub |>
  group_by(dataset_name) |>
  summarise(n = n(), 
            `Mean Age` = mean(age), 
            `Min Age` = min(age), 
            `Max Age` = max(age), 
            `CDIs` = ifelse(!is.na(any(prod)), "x", ""), 
            `longitudinal` = ifelse(any(duplicated(subject_id)), "x", "")) |>
  xtable(caption = "Characteristics of included datasets from Peekbank.")
```

```{r}
a <- ggplot(d_sub_long_elogit |> filter(measure == "long_window_elogit"), 
       aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_sub_agevar_long_elogit |> filter(measure == "long_window_elogit"),
              aes(group = dataset_name, col = dataset_name), geom="line",
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() +
  scale_color_discrete(guide = FALSE) + 
  ylab("Accuracy (elogit)") + 
  xlab("Age (months)")

b <- ggplot(d_sub_long_rt |> filter(measure == "log_rt"), 
       aes(x = age, y = rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_sub_agevar_long_rt |> filter(measure == "log_rt"),
              aes(group = dataset_name, col = dataset_name), geom="line",
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  # geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() +
  scale_color_discrete(guide = FALSE) + 
  theme(legend.position = "bottom") + 
  ylab("Reaction Time (log)") + 
  xlab("Age (months)")

f <- cowplot::plot_grid(a,b)
ggsave(f, filename = "figures/acc_rt.png", width = 8, height = 3.5)

>>>>

TODO:
- number of trials RT cleaning for slope comparison? (25% dataset? 2 RTs)
- check factor invariance across age - Measurement invariance?!?
- report on non-linearity model comparison

@@

```{r}
ggplot(d_sub_agevar, aes(x = age, y = long_window_accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

@@

kable(acc_mods_sub_lmer_summary, digits = 4)
```
With fixed effects.

```{r}
acc_mods_sub_lmer_fe <- list(lwe_lin = lmer(long_window_elogit ~ age_s  
                                         + (age_s | dataset_name), data = d_sub),
                          lwe_log = lmer(long_window_elogit ~ log_age_s 
                                         + (log_age_s | dataset_name), data = d_sub),
                          swe_lin = lmer(short_window_elogit ~ age_s 
                                         + (age_s | dataset_name), data = d_sub),
                          swe_log = lmer(short_window_elogit ~ log_age_s 
                                         + (log_age_s | dataset_name), data = d_sub),
                          lwa_lin = lmer(long_window_accuracy ~ age_s
                                         + (age_s | dataset_name), data = d_sub),
                          lwa_log = lmer(long_window_accuracy ~ log_age_s
                                         + (log_age_s | dataset_name), data = d_sub),
                          swa_lin = lmer(short_window_accuracy ~ age_s
                                         + (age_s | dataset_name), data = d_sub),
                          swa_log = lmer(short_window_accuracy ~ log_age_s
                                         + (log_age_s | dataset_name), data = d_sub))

acc_mods_sub_lmer_summary_fe <- map_df(acc_mods_sub_lmer_fe, ~broom.mixed::glance(.x)) |>
  mutate(model = names(acc_mods_sub_lmer_fe), 
         r2 = map_dbl(acc_mods_sub_lmer_fe, 
                      ~performance::r2_nakagawa(.x)$R2_conditional))

kable(acc_mods_sub_lmer_summary_fe, digits = 4)
```

### Reaction time models.

@@

```{r}
rt_mods_lmer_fe <- list(log_log = lmer(log_rt ~ log_age_s + 
                                      (log_age_s | dataset_name), data = d_sub),
                     log_lin = lmer(log_rt ~ age_s + 
                                      (age_s | dataset_name), data = d_sub),
                     lin_log = lmer(rt ~ log_age_s + 
                                      (log_age_s | dataset_name), data = d_sub),
                     lin_lin = lmer(rt ~ age_s + 
                                      (age_s | dataset_name), data = d_sub))

rt_mods_lmer_summary_fe <- map_df(rt_mods_lmer_fe, ~broom.mixed::glance(.x)) |>
  mutate(model = names(rt_mods_lmer_fe), 
         r2 = map_dbl(rt_mods_lmer_fe, ~performance::r2_nakagawa(.x)$R2_conditional))

kable(rt_mods_lmer_summary_fe, digits = 4)
```

### 2024-10-04 3_developmental_relationships.qmd
refactor

```{r}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(source(here("helper","common.R")))
library(lavaan)
library(tidySEM)

@@

```{r}
datasets_with_age_variation <- d_trial |>
  group_by(dataset_name) |>
  summarise(age_range = max(age) - min(age)) |>
  filter(age_range >= 4)

@@

```{r}
d_sub_mat <- d_sub |>
  ungroup() |>
  select(log_age, log_rt, long_window_accuracy)

@@

```{r}
mod1 <- fa(d_sub_mat, 1, rotate = "varimax")
mod1

@@

d_cfa <- d_sub |>
             rename(`accuracy` = long_window_accuracy) |>
             mutate(accuracy = scale(accuracy), 
                    log_age = scale(log_age), 
                    log_rt = scale(log_rt)) |>
  select(accuracy, log_age, log_rt)

@@

fa_model <-  "F1  =~ log_age + log_rt + accuracy"
fit <- cfa(fa_model, d_cfa, std.lv=TRUE)

summary(fit)
a <- graph_sem(model = fit)+ theme(panel.background = element_rect(fill = "white"))
```
Production.

@@

## Measurement invariance across age?

```{r}
fa_model <-  "F1  =~ log_age + log_rt + accuracy"
fit_younger <- cfa(fa_model, 
                   filter(d_cfa, log_age <= median(log_age)),
                   std.lv=TRUE)

graph_sem(model = fit_younger)
```

>>>>

```{r}
AGEVAR_CUTOFF <- 6

suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(source(here("helper","common.R")))
library(lavaan)
library(tidySEM)

@@

```{r}
datasets_with_age_variation <- d_trial |>
  group_by(dataset_name) |>
  summarise(age_range = max(age) - min(age)) |>
  filter(age_range >= AGEVAR_CUTOFF)

@@

```{r}
d_sub_mat <- d_sub |>
  ungroup() |>
  select(log_age, rt, long_window_accuracy)

@@

Now with variance too.

```{r}
d_sub_mat <- d_sub |>
  ungroup() |>
  select(log_age, rt, long_window_accuracy, long_window_acc_var, rt_var)

```{r}
mod1 <- fa(d_sub_mat, 1, rotate = "varimax")
mod1
```

@@

d_cfa <- d_sub |>
  rename(`accuracy` = long_window_accuracy) |>
  mutate(accuracy = scale(accuracy), 
         log_age = scale(log_age), 
         rt = scale(rt)) |>
  select(accuracy, log_age, rt)

@@

fa_model <-  "F1  =~ log_age + rt + accuracy"
fit <- cfa(fa_model, d_cfa, std.lv=TRUE)

summary(fit)
graph_sem(model = fit)+ theme(panel.background = element_rect(fill = "white"))
```
Production.

@@

d_cfa_dataset <- d_cfa |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name) |>
  group_by(dataset_name) |>
  nest() |>
  mutate(complete_data = map(data, ~filter(.x[complete.cases(.x),])),
         mat = as.matrix(complete_data), 
         has_data = map_lgl(complete_data, ~nrow(.x)>0)) |>
  filter(has_data) |>
  mutate(fa = map(mat, ~cfa(fa_model, .x, std.lv=TRUE)), 
         loadings = map(fa, ~summary(.x)[[5]]))

d_loadings <- d_cfa_dataset |>
  select(dataset_name, loadings) |>
  unnest(cols = c(loadings)) |>
  filter(!is.na(se))

ggplot(filter(d_loadings, lhs == "F1", rhs != "F1"),
       aes(x = rhs, y = est, col = dataset_name)) + 
  geom_point(alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot", aes(group = 1)) +
  geom_hline(yintercept = 0, lty = 2) + 
  coord_flip()+ 
  ylab("Standardized loading") + 
  xlab("Predictor") +
  scale_color_discrete(name = "Dataset")
```
With production data.

```{r}
fa_model_prod <-  "F1  =~ log_age + rt + accuracy + prod"

d_cfa_dataset_prod <-  d_sub |>
  rename(`accuracy` = long_window_accuracy) |>
  mutate(accuracy = scale(accuracy), 
         log_age = scale(log_age), 
         rt = scale(rt), 
         prod = scale(prod)) |>
  select(dataset_name, accuracy, log_age, rt, prod) |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name) |>
  group_by(dataset_name) |>
  nest() |>
  mutate(complete_data = map(data, ~filter(.x[complete.cases(.x),])),
         mat = as.matrix(complete_data), 
         has_data = map_lgl(complete_data, ~nrow(.x)>0)) |>
  filter(has_data) |>
  mutate(fa = map(mat, ~cfa(fa_model_prod, .x, std.lv=TRUE)), 
         loadings = map(fa, ~summary(.x)[[5]]))

d_loadings_prod <- d_cfa_dataset_prod |>
  select(dataset_name, loadings) |>
  unnest(cols = c(loadings)) |>
  filter(!is.na(se))

ggplot(filter(d_loadings_prod, lhs == "F1", rhs != "F1"),
       aes(x = rhs, y = est, col = dataset_name)) + 
  geom_point(alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot", aes(group = 1)) +
  geom_hline(yintercept = 0, lty = 2) + 
  coord_flip()+ 
  ylab("Standardized loading") + 
  xlab("Predictor") +
  scale_color_discrete(name = "Dataset")
```
## Measurement invariance across age?

This doesn't work.

```{r}
d_cfa_younger <- d_sub |>
    filter(age < median(age)) |>
  rename(`accuracy` = long_window_accuracy) |>
  mutate(accuracy = scale(accuracy), 
         log_age = scale(log_age), 
         log_rt = scale(log_rt)) |>
  select(accuracy, log_age, log_rt) 
d_cfa_younger <- d_cfa_younger[complete.cases(d_cfa_younger), ]

d_cfa_older <- d_sub |>
  filter(age >= median(age)) |>
  rename(`accuracy` = long_window_accuracy) |>
  mutate(accuracy = scale(accuracy), 
         log_age = scale(log_age), 
         log_rt = scale(log_rt)) |>
  select(accuracy, log_age, log_rt) 
d_cfa_older <- d_cfa_older[complete.cases(d_cfa_older), ]

fit_younger <- cfa(fa_model, 
                   data = d_cfa_younger,
                   std.lv=TRUE)
fit_older <- cfa(fa_model, 
                 data = d_cfa_older,
                 std.lv=TRUE)

younger_plot <- graph_sem(model = fit_younger)
older_plot <- graph_sem(model = fit_older)
plot_grid(younger_plot, older_plot)
```
